---
title: "Praca domowa 1"
author: "Martyna Majchrzak"
date: "9 03 2020"
output: html_document
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# biblioteki
library(DataExplorer)
library(dplyr)
library(ggplot2)

#załadowanie i przygotowanie danych
data<-read.csv("../data/heart_disease_weka_dataset.csv")

```
## Dataset

Information about patients from the Cleveland hospital being diagnosed with heart disease.

Columns:


- age	- age of patient

- sex	-	1=male; 0=female

- cp	-	chest pain type: 1=typical angina; 2=atypical angine;  3=non-anginal pain; 4=asymptomatic

- trestbps	-	resting blood pressure (mm Hg)

- chol	-	serum cholestrol (mg/dl)

- fbs	-	fasting blood sugar: 1 if > 120 mg/dl; 0 otherwise

- restecg	-	resting electrocardiographic results: 0=normal; 1=having ST-T wave abnormality; 2=showing probable or definite left ventricular hypertrophy

- thalach	-	maximum heart rate achieved

- exang	-	exercise induced angina: 1=yes; 0=no

- oldpeak	-	ST depression induced by exercise relative to rest

- slope	-	the slope of the peak exercise ST segment: 1=upsloping; 2=flat; 3=downsloping

- ca	-	number of major vessels (0-3) colored by flourosopy

- thal	-	3=normal; 6=fixed defect; 7=reversable defect

- [target] num - 0-HEART DISEASE, 1-NO HEART DISEASE)


```{r}
head(data)
summary(data)
```

## Data Explorer

#### Density 
```{r}
DataExplorer::plot_density(data, title="All")
DataExplorer::plot_density(data%>%filter(num==0), title="Heart disease")
DataExplorer::plot_density(data%>%filter(num==1), title="No heart disease")

```

Patients with heart disease are much more common to have non-anginal chest pain (3), where patients without hearts disease often have asymptomatic chest pain (4).

#### Boxplots

```{r}
DataExplorer::plot_boxplot(data, by = "num")
```

Patients with heart disease have slighty maximum heart rate achieved (thalach) than those without.

```{r}
DataExplorer::plot_boxplot(data, by = "sex")
```

Women have slighty higher maximum heart rate achieved (thalach) and cholesterol (ch) than man, 
```{r}
DataExplorer::plot_boxplot(data, by = "age")
```

Resting blood pressure (trestbps) is increasing with age, but maximum heart rate (thalach) dicreases.
```{r}
DataExplorer::plot_qq(data)
```

#### Correlation

```{r}
DataExplorer::plot_correlation(data[,-1])
```
