---

title: "SVM"

author: "Mikołaj Malec"

date: "4/28/2020"

output: html_document

---



## Data *apartments* from DALEX



```{r main, echo=TRUE, message=TRUE, warning=FALSE}
set.seed(123)

df_apartments <- DALEX::apartments

#split

n_train <- 1000

n_split <- n_train * 9/10

ind_sam <- sample( 1:dim( DALEX::apartments)[1], n_train)



df_train <- df_apartments[ind_sam[1:n_split],]

df_test <- df_apartments[ind_sam[(n_split+1):n_train],]

## explore data

# first rows

head( df_apartments)

# summary of columns

summary( df_apartments)

# dimensions of data frame and types of columns

str( df_apartments)

# there is no missing data

sum( is.na( df_apartments))



# distribution of target

plot( df_train[6], las=2)

```



Because target is balanced, that means that accuracy is the best metric for this example.



### Random search of hiperparameters for SVM



Color represents a degree parameter.

Click plot for interactivity.


```{r apa train, echo=FALSE, message=FALSE, warning=FALSE}

#metric

accuracy <- function( tab){

  sum( diag(tab)) / sum( tab)

}





## training

library( e1071)



#random search


n <- 100

C <- 10 ** runif( n, -3, 2)

gamma <- 10 ** runif( n, -3, 0) #hogher gamma lead to too many iteracions

degree <- sample( c(2,3,4), n, replace = TRUE)



table_svm_apatments_list <- list()

accuracy_vec <- rep( 0, n)



for( i in 1:n){

  model_svm_apatments <- svm( district~., data = df_train,

                              kernel = "polynomial",

                              degree = degree[i],

                              cost = C[i],

                              gamma = gamma[i])

  pred_svm_apatments <- predict( model_svm_apatments, df_test[-6])

  table_svm_apatments <- table( pred_svm_apatments, df_test[,6])

  

  table_svm_apatments_list[[i]] <- table_svm_apatments

  accuracy_vec[i] <- accuracy( table_svm_apatments)

}



imax <- which.max( accuracy_vec)



degree <- as.character( degree)



out <- data.frame( C, gamma, degree, accuracy_vec)



library( plotly)

p <- plot_ly( data = out, x = ~C, y= ~gamma, 

              color = ~degree, size = ~accuracy_vec, text = accuracy_vec,

              hovertemplate = paste("C: %{x}",

                                    "<br>gamma: %{y}",

                                    "<br>Accuracy: %{text}"))

p <- add_annotations( p,

                 x = log10(out$C[imax]),

                 y = log10(out$gamma[imax]),

                 text = "Best accuracy")

p <- layout(p, xaxis = list(type = "log"),

            yaxis = list(type = "log"))

p <- add_markers(p)

p

```



Best accuracy = `r accuracy_vec[imax]` is for C = `r C[imax]`, gamma = `r gamma[imax]` and  degree = `r degree[imax]`



Data has been randomly generated. Many districts aren"t diverse to be sliced. Exeption is Srodmiescie which is very distinguishable for others. It"s no worry that accuracy is so small. Confusion matrix for the best is:



```{r table, echo=FALSE}

library(knitr)

knitr::kable(table_svm_apatments_list[[imax]])

```



## Data *weather* from nycflights13



```{r main 2, echo=TRUE, message=TRUE, warning=FALSE}

#delete columns with many missing

DataExplorer::profile_missing( nycflights13::weather)

df_weather <- nycflights13::weather[,c(-9,-11,-13)]

#delete rows with some missing obesvations

df_weather <- df_weather[ complete.cases(df_weather),]



#no need for posic column and date columns

df_weather$time_hour <- NULL

df_weather$year <- NULL

df_weather$month <- NULL

df_weather$day <- NULL

df_weather$hour <- NULL



df_weather$origin <- as.factor( df_weather$origin)



#split

n_train <- 1000

n_split <- n_train * 9/10

ind_sam <- sample( 1:dim(nycflights13::weather)[1], n_train)



df_train <- df_weather[ind_sam[1:n_split],]

df_test <- df_weather[ind_sam[(n_split+1):n_train],]



## explore data

# first rows

head( df_weather)

# summary of columns

summary( df_weather)

# dimensions of data frame and types of columns

str( df_weather)

# there is no longer missing data

sum( is.na( df_weather))



# distribution of target

knitr::kable( table(df_train["origin"]))

```



Because target is balanced, that means that accuracy is the best metric for this example.



### Random search of hiperparameters for SVM



Color represents a degree parameter.

Click plot for interactivity.

```{r weather train, echo=FALSE, message=FALSE, warning=FALSE}

#metric

accuracy <- function( tab){

  sum( diag(tab)) / sum( tab)

}





#random search

n <- 100

C <- 10 ** runif( n, -3, 2)

gamma <- 10 ** runif( n, -3, 0) #higher gamma lead to too many iterations

degree <- sample( c(2,3,4), n, replace = TRUE)



table_svm_weather_list <- list()

accuracy_vec <- rep( 0, n)



for( i in 1:n){

  model_svm_weather <- svm( origin~., data = df_train,

                              kernel = "polynomial",

                              degree = degree[i],

                              cost = C[i],

                              gamma = gamma[i])

  #origin =1

  pred_svm_weather <- predict( model_svm_weather, df_test[-1])

  table_svm_weather <- table( pred_svm_weather, unlist(df_test[,1]))

  

  table_svm_weather_list[[i]] <- table_svm_weather

  accuracy_vec[i] <- accuracy( table_svm_weather)

}

imax <- which.max( accuracy_vec)



degree <- as.character( degree)



out <- data.frame( C, gamma, degree, accuracy_vec)



library( plotly)

p <- plot_ly( data = out, x = ~C, y= ~gamma, 

              color = ~degree, size = ~accuracy_vec, text = accuracy_vec,

              hovertemplate = paste("C: %{x}",

                                    "<br>gamma: %{y}",

                                    "<br>Accuracy: %{text}"))

p <- add_annotations( p,

                 x = log10(out$C[imax]),

                 y = log10(out$gamma[imax]),

                 text = "Best accuracy")

p <- layout(p, xaxis = list(type = "log"),

            yaxis = list(type = "log"))

p <- add_markers(p)

p

```



Best accuracy = `r accuracy_vec[imax]` is for C = `r C[imax]`, gamma = `r gamma[imax]` and  degree = `r degree[imax]`



As an experiment I have chosen weather data knowing, that the results will be poor, especially when categories are very similar. My idea was to check is SVM any good in overlapping data. It isn't. But the second part of my little experiment is interesting. Hiperparameters behaved very differently than before. Before, top right corner had the best parameters. Now is opposite. Top right has occurence of the worst cases. Confusion matrix for the best is:



```{r table 2, echo=FALSE}

library(knitr)

knitr::kable(table_svm_weather_list[[imax]])

```



```{r info}

sessionInfo()

```